Self Regulation Ontology Retest Data: Meaningful Variables Clean

### 2017-03-07 SRO_Meaningful_Vars.Rmd
made new notebook just looking at meaningful vars

---
title: 'Self Regulation Ontology Retest Data: Initial Exploration'
output:
html_document:
toc: yes
toc_float: yes
---

```{r, message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(lme4)
library(GGally)
library(jsonlite)
library(lavaan)
library(semTools)
library(psych)
library(GPArotation)
library(rmarkdown)
sem <- function(x) {sd(x, na.rm=T) / sqrt(length(x))}
render_this <- function(){rmarkdown::render('SRO_Retest_Analyses.Rmd', html_notebook(toc = T, toc_float = T))}
options(digits = 4)
```

@@

### Load original data
```{r}
test_data <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Complete_01-31-2017/variables_exhaustive.csv')
```

### Load retest data
```{r}
retest_data <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Retest_02-11-2017/variables_exhaustive.csv')
```

@@

Drop columns from original data that are not in retest data

Are there any columns in retest data columns that are not in the original data: No.

```{r}
names(retest_data)[names(retest_data) %in% names(retest_subs_test_data) == FALSE]
```

Test data columns that are not in the retest data: Two stage hasn't been fixed yet.

```{r}
names(retest_subs_test_data)[names(retest_subs_test_data) %in% names(retest_data) == FALSE]
```

Datasets with only matching columns

retest_data <- retest_data[,matching_columns]
retest_subs_test_data <- retest_subs_test_data[,matching_columns]
```

# Correlations of exhaustive variables

Get list of test retest correlations for all matching (numeric) columns in the test and retest data. Note two caveats:  
**1. There are still some columns that weren't in both datasets listed above (e.g. two stage)**  
**2. Unlike meaningful variables these variables are not (yet) rid of outliers or log transformed if skewed. This will be detailed below along with its problems.**

```{r}
retest_data <- as.data.frame(retest_data)
retest_subs_test_data <- as.data.frame(retest_subs_test_data)

cor_df <- data.frame(pearson = rep(NA, length(matching_dv_columns)), spearman = rep(NA, length(matching_dv_columns)))

for(i in 1:length(matching_dv_columns)){
  cor_df[matching_dv_columns[i], 'pearson'] <- cor(retest_data[,matching_dv_columns[i]], retest_subs_test_data[,matching_dv_columns[i]], method = 'pearson', use = "pairwise.complete.obs")
  cor_df[matching_dv_columns[i], 'spearman'] <- cor(retest_data[,matching_dv_columns[i]],retest_subs_test_data[,matching_dv_columns[i]], method = 'spearman', use = "pairwise.complete.obs")
}
```

Distribution of all Pearson and Spearman correlations
```{r warning=FALSE, message=FALSE}
cor_df %>%
  gather(key, value) %>%
  ggplot(aes(value))+
  geom_histogram()+
  theme_bw()+
  facet_wrap(~key)
```

Survey reliabilities are noticably higher than task reliabilities.

Task reliability distributions looks alsmost bimodal. **Should figure out whether the lower ones are specific tasks or if they are certain kinds of vars (e.g. certain ddm parameters)**
```{r warning=FALSE, message=FALSE}
cor_df %>%
  gather(key, value, -dv, -task) %>%
  ggplot(aes(value))+
  geom_histogram()+
  theme_bw()+
  facet_wrap(key~task)
```
Mean reliability and the number of variables for both the surveys and the tasks
```{r}
cor_df %>%
  group_by(task) %>%
  summarise(mean_pearson = mean(pearson),
            mean_spearman = mean(spearman),
            num_vars = n())
```

#### For tasks
```{r}
cor_df %>%
  filter(task == 'task') %>%
  select(dv, spearman, pearson) %>%
  arrange(-pearson, -spearman)
```

Note: Holt and Laury test-retest is pretty terrible. In previous data I have worked with it was ~.35. Should check to see if there is a coding error there.
```{r}
cor_df %>%
  filter(task == 'survey') %>%
  select(dv, spearman, pearson) %>%
  arrange(-pearson, -spearman)
```

The smaller meaningful_vars dataframe has been somewhat cleaned. So far in this notebook we looked at all the vars in their rawest form. Here we'll first check the distributions of all the variables and then apply the same cleaning procedure that was used for the meaningful variables to all of them.

Note that I'm doing this for the sake of consistency for now. We might later decide that we would like to transform certain variables of interest in other ways to be able to include them in our analyses without violating the assumptions of those analyses. The current procedure checks the skewness of a column first removes the outliers defined as greater than 2.5 IQR from median then variables with skewness >1 are log transformed.

First I'll check if the variables that went in to the correlation matrix have normal distributions since at least the Pearson correlations depend on this assumption. Then everything will be cleaned the way meaningful variables was cleaned. There is a problem with this approach that I'm not sure how to deal with yet but this will be detailed below.

Function to check if the variables that went in to cor_df have normal distributions
```{r}
diff_from_normal <- function(col){
  return(shapiro.test(col)$p.value < 0.05)
}
```

Create and organize new df where distribution of vars data is stored. First for retest data only; then for test data for subjects who have retest data - restricting here since these are the subsets of data that went in to the correlation matrix. Merge and cleanup.
```{r}
dist_retest_data <- as.data.frame(apply(retest_data[,matching_dv_columns], 2, diff_from_normal))
names(dist_retest_data) <- c('retest_not_normal')
dist_retest_data$dv <- row.names(dist_retest_data)
row.names(dist_retest_data) <- seq(1:nrow(dist_retest_data))
dist_retest_data = dist_retest_data %>% select(dv, retest_not_normal)

rm(dist_retest_data, dist_retest_subs_test_data)
```

Add whether dv's are tasks or survey to distribution dfs
```{r}
dist_data = cor_df %>%
  select(dv, task) %>%
  left_join(dist_data, by = 'dv')
```

**Are the same vars that are not normally distributed in the retest data not normal in test data as well?**  
List of variables that have different distributions in the two time points.
A majority of them are drift diffusion parameters (but note that these vars are overrepresented to begin with - more on them later).
```{r}
dist_data %>%
  filter(retest_not_normal != test_not_normal)
```

Apply the same cleaning procedure that was used for meaningful vars for all vars: 
```{r}
remove_outliers = function(data_column, quantile_range = 2.5){

@@

"%w/o%" <- function(x, y) x[!x %in% y]

neg_log <- function(column){
  col_max = max(column, na.rm=T)
  column = col_max+1-column
  return(log(column))
}

transform_remove_skew = function(data, columns, threshold = 1){

  tmp = as.data.frame(apply(data[,columns],2,skew))
  names(tmp) = c("skew")
  tmp$dv = row.names(tmp)
  tmp = tmp %>% 
    filter(abs(skew)>threshold)

  skewed_variables = tmp$dv
  skew_subset = data[, skewed_variables]
  positive_subset = data[,tmp$dv[tmp$skew>0]]
  negative_subset = data[,tmp$dv[tmp$skew<0]]

  # transform variables
  # log transform for positive skew
  positive_subset = log(positive_subset)
  successful_transforms = as.data.frame(apply(positive_subset, 2, skew))
  names(successful_transforms) = c('skew')
  successful_transforms$dv = row.names(successful_transforms)
  successful_transforms = successful_transforms %>% filter(abs(skew)<threshold)
  successful_transforms = successful_transforms$dv
  successful_transforms = positive_subset[,successful_transforms]
  dropped_vars = names(positive_subset) %w/o% names(successful_transforms)

  # replace transformed variables
  data = data[,-c(which(names(data) %in% names(positive_subset)))]
  names(successful_transforms) = paste0(names(successful_transforms), '.logTr')
  cat(rep('*', 40))
  cat('\n')
  cat(paste0('Dropping ', length(dropped_vars) ,' positively skewed data that could not be transformed successfully:'))
  cat('\n')
  cat(dropped_vars, sep = '\n')
  cat(rep('*', 40))
  cat('\n')
  data = cbind(data, successful_transforms)

  # reflected log transform for negative skew      
  negative_subset = as.data.frame(apply(negative_subset, 2, neg_log))
  successful_transforms = as.data.frame(apply(negative_subset, 2, skew))
  names(successful_transforms) = c('skew')
  successful_transforms$dv = row.names(successful_transforms)
  successful_transforms = successful_transforms %>% filter(abs(skew)<1)
  successful_transforms = successful_transforms$dv
  successful_transforms = negative_subset[,successful_transforms]
  dropped_vars = names(negative_subset) %w/o% names(successful_transforms)

  # replace transformed variables
  data = data[,-c(which(names(data) %in% names(negative_subset)))]
  names(successful_transforms) = paste0(names(successful_transforms), '.ReflogTr')
  cat(rep('*', 40))
  cat('\n')
  cat(paste0('Dropping ', length(dropped_vars) ,' negatively skewed data that could not be transformed successfully:'))
  cat('\n')
  cat(dropped_vars, sep = '\n')
  cat(rep('*', 40))
  cat('\n')
  data = cbind(data, successful_transforms)

  return(data)
}
```

Clean retest data with these rules.
```{r warning=FALSE}
clean_retest_data = as.data.frame(apply(retest_data[, matching_dv_columns], 2, remove_outliers))
clean_retest_data = transform_remove_skew(clean_retest_data, matching_dv_columns)
```

**I'm not sure what the right thing to do is regarding the cleaning of the test data for the retest subjects.**   
Should they be cleaned on the whole sample for the time they were collected on and then extracted (this should have an effect on outliers as well as transformations) or extracted and then cleaned. Here I am doing the latter but this seems to be resulting in more skewed variables for test data compared to the retest data so the later correlation matrix might not include all the variables we would like to look at.

```{r warning=FALSE}
clean_retest_subs_test_data = as.data.frame(apply(retest_subs_test_data[, matching_dv_columns], 2, remove_outliers))
clean_retest_subs_test_data = transform_remove_skew(clean_retest_subs_test_data, matching_dv_columns)
```

## Correlations on cleaned variables

Create the same list of correlations as before but this time on a smaller set of cleaned variables.

```{r warning=FALSE, message=FALSE}
clean_all_columns <- unique(c(names(clean_retest_data), names(clean_retest_subs_test_data)))

clean_matching_columns <- c()
for(i in 1:length(clean_all_columns)){
  if(clean_all_columns[i] %in% names(clean_retest_data) & clean_all_columns[i] %in% names(clean_retest_subs_test_data)){
    clean_matching_columns <- c(clean_matching_columns, clean_all_columns[i])
  }
}

# clean_retest_data <- clean_retest_data[,clean_matching_columns]
# clean_retest_subs_test_data <- clean_retest_subs_test_data[,clean_matching_columns]

clean_matching_dv_columns <- c()

for(i in 1:length(clean_matching_columns)){
  if(is.numeric(clean_retest_data[,clean_matching_columns[i]]) & is.numeric(clean_retest_subs_test_data[,clean_matching_columns[i]])){
    clean_matching_dv_columns <- c(clean_matching_dv_columns, clean_matching_columns[i])
  }
}

clean_cor_df <- data.frame(pearson = rep(NA, length(clean_matching_dv_columns)), spearman = rep(NA, length(clean_matching_dv_columns)))

row.names(clean_cor_df) <- clean_matching_dv_columns

for(i in 1:length(clean_matching_dv_columns)){

  robustCor <- function(...) {
    obj<-try(cor(...), silent=TRUE)
    if (is(obj, "try-error")) return(NA) else return(obj)
  }

  tmp_pearson = robustCor(clean_retest_data[,clean_matching_dv_columns[i]], clean_retest_subs_test_data[,clean_matching_dv_columns[i]], method = 'pearson', use = "pairwise.complete.obs")

  tmp_spearman = robustCor(clean_retest_data[,clean_matching_dv_columns[i]], clean_retest_subs_test_data[,clean_matching_dv_columns[i]], method = 'spearman', use = "pairwise.complete.obs")

  clean_cor_df[clean_matching_dv_columns[i], 'pearson'] <- tmp_pearson 
  clean_cor_df[clean_matching_dv_columns[i], 'spearman'] <- tmp_spearman

}

clean_cor_df$dv = row.names(clean_cor_df)
row.names(clean_cor_df) = seq(1:nrow(clean_cor_df))
clean_cor_df$task = 'task'
clean_cor_df[grep('survey', clean_cor_df$dv), 'task'] = 'survey'
```

Correlations could not be calculated for the following variables due to lack of variance.

```{r}
clean_cor_df %>%
  filter(is.na(pearson)) %>%
  select(dv)
```

Let's look at the distributions again: The distributions don't look terribly different but there are noticably fewer variables post cleaning. 
```{r warning=FALSE, message=FALSE}
cor_df %>%
  left_join(clean_cor_df, by='dv') %>%
  select(-task.y)%>%
  rename(pearson_preClean = pearson.x, spearman_preClean = spearman.x,
         pearson_postClean = pearson.y, spearman_postClean = spearman.y,
         task = task.x) %>%
  gather(key, value, -dv, -task) %>%
  separate(key, c('cor_type', 'clean_type'),sep = "_") %>%
  ggplot(aes(x=value, fill=factor(clean_type, levels = c('preClean', 'postClean'),
                                  labels = c('Before', 'After'))))+
  geom_histogram(position='identity', alpha = 0.5)+
  facet_wrap(cor_type~task)+
  theme_bw()+
  theme(legend.title = element_blank())
```

This cleaning led to a slight change in the mean reliabilities but a lot of dropped variables.
```{r}
clean_cor_df %>%
  group_by(task) %>%
  summarise(mean_pearson = mean(pearson, na.rm=T),
            mean_spearman = mean(spearman, na.rm=T),
            num_vars = sum(ifelse(is.na(pearson)==FALSE,1,0)))
```

**What changed most with cleaning?**    
Difference in correlation pre and post correlation for the variables that survived cleaning.  
Things that changed most seem to be RTs and parameter estimates.

```{r}
clean_cor_df %>%
  left_join(cor_df, by='dv') %>%
  mutate(pearson_diff = abs(pearson.y - pearson.x),
         spearman_diff = abs(spearman.y - spearman.x)) %>%
  select(dv, pearson_diff, spearman_diff) %>%
  arrange(-pearson_diff, -spearman_diff)
```

@@

We should probably take a closer look at model fits instead but this could be a first step.  
I'm not sure why there are so many more EZ variables than hddm variables.

```{r message=FALSE, warning=FALSE, fig.height=12, fig.width=9}
cor_df%>%
  select(-pearson) %>%
  filter(grepl('EZ|hddm|_rt|_acc', dv)) %>%
  mutate(var_type = ifelse(grepl('EZ',dv), 'EZ', ifelse(grepl('hddm',dv), 'hddm', 'raw')),
         var = ifelse(grepl('_rt', dv), 'rt', ifelse(grepl('_acc', dv), 'acc', ifelse(grepl('_non_decision', dv), 'non_decision', ifelse(grepl('_thresh', dv), 'thresh', ifelse(grepl('_drift', dv), 'drift', NA)))))) %>%
  left_join(clean_cor_df[,c('spearman', 'dv')], by = 'dv') %>%
  gather(clean_type, spearman, -dv, -var_type, -var, -task) %>%
  mutate(clean_type = ifelse(clean_type == 'spearman.x','Before','After')) %>%
  ggplot(aes(spearman, fill = clean_type))+
  geom_histogram(position='identity', alpha = 0.5)+
  theme_bw()+
  facet_wrap(var_type~var)+
  scale_fill_discrete(name = 'Clean type')

Average reliabilities

1. Cleaning doesn't seem to make a huge difference (at least for Spearman correlations) except for raw accuracies and hddm non_decision  
2. There are many more EZ variabiles but the few hddm variables have slightly higher reliabilities (I'm not sure about the reason for the discrepancy between the number of variables)  
3. Raw accuracies seem to have the worst reliabilities on average while hddm parameters the highest

```{r}
cor_df%>%
  select(-pearson) %>%
  filter(grepl('EZ|hddm|_rt|_acc', dv)) %>%
  mutate(var_type = ifelse(grepl('EZ',dv), 'EZ', ifelse(grepl('hddm',dv), 'hddm', 'raw')),
         var = ifelse(grepl('_rt', dv), 'rt', ifelse(grepl('_acc', dv), 'acc', ifelse(grepl('_non_decision', dv), 'non_decision', ifelse(grepl('_thresh', dv), 'thresh', ifelse(grepl('_drift', dv), 'drift', NA)))))) %>%
  left_join(clean_cor_df[,c('spearman', 'dv')], by = 'dv') %>%
  gather(clean_type, spearman, -dv, -var_type, -var, -task) %>%
  mutate(clean_type = ifelse(clean_type == 'spearman.x','Before','After')) %>%
  group_by(var_type, clean_type, var) %>%
  summarise(mean_spearman = mean(spearman, na.rm=T),
            sem_spear = sem(spearman),
            num_vars = sum(ifelse(is.na(spearman)==FALSE,1,0))) %>%
  arrange(var_type, -as.numeric(as.factor(clean_type)))
```

## Other variable types

I planned to look at differences between distributions of reliabilities for other variables types as well but I am not sure what these categories should be so I am deferring this for further discussions. My earlier notes suggested differences between other model parameters versus raw variables; subtraction versus basic variables; single versus multiple variables.I'm hoping we can come up with a more systematic definition that I can automatize or at least restrict to certain groups of variable instead of going through all 600+ variables and manually labeling them.

@@

First do a t-test on all measures of T1 for subjects with and without retest to confirm that these two samples do not differ from each other significantly
```{r}
ttest_df = data.frame(t_stat = rep(NA, length(matching_dv_columns)),
                      p_val = rep(NA, length(matching_dv_columns)))

@@

  ttest_df[matching_dv_columns[i], 'p_val'] <- t.test(retest_subs_test_data[,matching_dv_columns[i]], rest_subs_test_data[,matching_dv_columns[i]])$p.value
}

ttest_df$dv = row.names(ttest_df)
row.names(ttest_df) = seq(1:nrow(ttest_df))
ttest_df$task = 'task'
ttest_df[grep('survey', ttest_df$dv), 'task'] = 'survey'
```

Distribution of t values

```{r warning=FALSE, message=FALSE}
ttest_df %>%
  ggplot(aes(t_stat)) +
  geom_histogram()+
  theme_bw()+
  facet_wrap(~task)
```

@@

43 variables. None of them are very strongly different though.

```{r}
ttest_df %>%
  filter(p_val<0.05) %>%
  select(dv, task, t_stat, p_val)
```

Do this again with cleaned data.  
Clean the T1 data for the subsets first.  
**Here again I'm not sure if I should clean before or after splitting. Before it seemed appropriate to clean subjects with retest data after separating them from the remaining subjects but here it might be better to clean first and then separate. But then it would differ from what was done for earlier anaylses and it seems like the kind of thing that should be kept consistent.**

Clean T1 for subjects with retest data (doing it again because this df was subsetted before)

```{r warning=FALSE}
clean_retest_subs_test_data = as.data.frame(apply(retest_subs_test_data[, matching_dv_columns], 2, remove_outliers))
clean_retest_subs_test_data = transform_remove_skew(clean_retest_subs_test_data, matching_dv_columns)
```

Clean T1 for rest of the subjects

```{r warning=FALSE}
clean_rest_subs_test_data = as.data.frame(apply(rest_subs_test_data[, matching_dv_columns], 2, remove_outliers))
clean_rest_subs_test_data = transform_remove_skew(clean_rest_subs_test_data, matching_dv_columns)
```

Get numeric dv's from both cleaned datasets to compute t-tests on.
```{r}
clean_all_columns <- unique(c(names(clean_rest_subs_test_data), names(clean_retest_subs_test_data)))

clean_matching_columns <- c()
for(i in 1:length(clean_all_columns)){
  if(clean_all_columns[i] %in% names(clean_rest_subs_test_data) & clean_all_columns[i] %in% names(clean_retest_subs_test_data)){
    clean_matching_columns <- c(clean_matching_columns, clean_all_columns[i])
  }
}

clean_matching_dv_columns <- c()

for(i in 1:length(clean_matching_columns)){
  if(is.numeric(clean_rest_subs_test_data[,clean_matching_columns[i]]) & is.numeric(clean_rest_subs_test_data[,clean_matching_columns[i]])){
    clean_matching_dv_columns <- c(clean_matching_dv_columns, clean_matching_columns[i])
  }
}
```

Compute t-tests for cleaned variables.
```{r}
clean_ttest_df = data.frame(t_stat = rep(NA, length(clean_matching_dv_columns)),
                      p_val = rep(NA, length(clean_matching_dv_columns)))

row.names(clean_ttest_df) <- clean_matching_dv_columns

for(i in 1:length(clean_matching_dv_columns)){

  robustTTest <- function(...) {
    obj<-try(t.test(...), silent=TRUE)
    if (is(obj, "try-error")) return(obj = data.frame(statistic = NA, p.value=NA)) else return(obj)
  }

  tmp = robustTTest(clean_retest_subs_test_data[,clean_matching_dv_columns[i]], clean_rest_subs_test_data[,clean_matching_dv_columns[i]])

  clean_ttest_df[clean_matching_dv_columns[i], 't_stat'] <- tmp$statistic 
  clean_ttest_df[clean_matching_dv_columns[i], 'p_val'] <- tmp$p.value
    }

clean_ttest_df$dv = row.names(clean_ttest_df)
row.names(clean_ttest_df) = seq(1:nrow(clean_ttest_df))
clean_ttest_df$task = 'task'
clean_ttest_df[grep('survey', clean_ttest_df$dv), 'task'] = 'survey'
```

Variables for which t-tests cannot be computed due to lack of variance.
```{r}
clean_ttest_df %>%
  filter(is.na(t_stat)) %>%
  select(dv)
```

Distribution of t-tests for the cleaned variables with sufficient variance
```{r warning=FALSE, message=FALSE}
clean_ttest_df %>%
  filter(!is.na(t_stat)) %>%
ggplot(aes(t_stat)) +
  geom_histogram()+
  theme_bw()+
  facet_wrap(~task)
```

**Which variables are significantly different between the two samples on cleaned data (those with and without retest data for T1)?**

More variables (that are thought to have been properly cleaned for these analyses) are significantly different between the two samples.   
**Not sure what to do with this especially for the next set of analyses.**

```{r}
clean_ttest_df %>%
  filter(!is.na(p_val) & p_val<0.05) %>%
  select(dv, task, t_stat, p_val)
```

@@

Step 2: Efa on the people you don't have retest on and then predict how well this holds up for people w retest on t1 and on t2.

I'm restricting the factor analyses to the meaningful variables for now (instead of the 600+ exhaustive variables) to keep it manageable. I haven't consulted other factor analyses that have been done on the full dataset in doing this but will be happy to adjust as necessary.

Data cleaning for efa
```{r eval=FALSE}
#read in only meaningful vars
meaningful_var_test <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Complete_01-31-2017/meaningful_variables_clean.csv')

meaningful_var_retest <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Retest_02-11-2017/meaningful_variables_clean.csv')

#Process sub_id columns in the data dataframes
meaningful_var_retest$X <- as.character(meaningful_var_retest$X)
meaningful_var_test$X <- as.character(meaningful_var_test$X)

names(meaningful_var_retest)[1] <- 'sub_id'
names(meaningful_var_test)[1] <- 'sub_id'

#Clean and subset dfs as before
meaningful_var_retest <- meaningful_var_retest %>%
  group_by(sub_id) %>%
  do(correct_subids(.,test_worker_lookup))

#Remove sub w missing test data from retest data
retest_data_mngf <- meaningful_var_retest[meaningful_var_retest$sub_id %in% meaningful_var_test$sub_id,]

#Extract test data for subs w retest data
retest_subs_test_mngf <- meaningful_var_test[meaningful_var_test$sub_id %in% meaningful_var_retest$sub_id,]

#Extract test data for subs without retest data (efa df)
rest_subs_test_mngf <- meaningful_var_test[meaningful_var_test$sub_id %in% meaningful_var_retest$sub_id == FALSE,]

#Arrange datasets of same size by sub_id
retest_data_mngf = retest_data_mngf %>% arrange(sub_id)
retest_subs_test_mngf = retest_subs_test_mngf %>% arrange(sub_id)

#CHECK IF EVERYTHING IS ORDERED RIGHT
retest_subs_test_mngf$sub_id == retest_data_mngf$sub_id
```

Checking columns for efa:
There
```{r eval=FALSE}
#Check columns 
names(retest_data_mngf)[names(retest_data_mngf) %in% names(retest_subs_test_mngf) == FALSE]

names(retest_subs_test_mngf)[names(retest_subs_test_mngf) %in% names(retest_data_mngf) == FALSE]
```

For now I'm just removing these variables but will have to do re-do them once I resolve why they weren't in both datasets to begin with (The answer will be a combination of not having run the most recent analysis script on some tasks and dropped variables as a result of cleaning)
```{r}
#Subset datasets with only matching columns
all_columns_mngf <- unique(c(names(retest_data_mngf), names(retest_subs_test_mngf)))

matching_columns_mngf <- c()
for(i in 1:length(all_columns_mngf)){
  if(all_columns_mngf[i] %in% names(retest_data_mngf) & all_columns_mngf[i] %in% names(retest_subs_test_mngf)){
    matching_columns_mngf <- c(matching_columns_mngf, all_columns_mngf[i])
  }
}

>>>>

---
title: 'Self Regulation Ontology Retest Data: Meaningful Variables Clean'
output:
html_document:
toc: yes
toc_float: yes
---

```{r, message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(lme4)
library(GGally)
library(jsonlite)
library(lavaan)
library(semTools)
library(psych)
library(GPArotation)
library(rmarkdown)
sem <- function(x) {sd(x, na.rm=T) / sqrt(length(x))}
render_this <- function(){rmarkdown::render('SRO_Meaningful_Vars.Rmd', html_notebook(toc = T, toc_float = T))}
options(digits = 4)
```

@@

### Load original data (only cleaned meaningful variables)
```{r}
test_data <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Complete_01-31-2017/meaningful_variables_clean.csv')
```

### Load retest data (all variables; will subset and clean based on cleaned variables for test data)
```{r}
retest_data <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Retest_02-11-2017/variables_exhaustive.csv')
```

@@

```{r echo=FALSE}
rm(test_worker_lookup, correct_subids, process_lookup)
```

Get only cleaned meaningful var columns based on test data from retest data

```{r}
#Get list of meaningful vars
meaningful_vars = names(test_data)

#Remove transofrmation suffices to be able to use for subsetting of retest data
meaningful_vars = gsub(".logTr|.ReflogTr", "", meaningful_vars)

#Get list of transformed variables and the sign of their skew
logTr_vars = data.frame(grep('logTr', names(test_data), value=T))
names(logTr_vars) = c('var')
logTr_vars$sign = ifelse(grepl(".ReflogTr",logTr_vars$var), 'negative', 'positive')
logTr_vars$var = gsub(".logTr|.ReflogTr", "", logTr_vars$var)

#Subset retest data
retest_data = retest_data[, names(retest_data) %in% meaningful_vars]
```

Clean columns subsetted from retest data the way they have been for test data 
**Question: Should I be removing outliers from the retest data? Or keep them if they were within the ranges in the test data even if they aren't in the retest data**
````{r}
#Function to remove outliers as they have been removed for the test data
remove_outliers = function(data_column, quantile_range = 2.5){

@@

#Function to transform the variables with negative skew
neg_log <- function(column){
  col_max = max(column, na.rm=T)
  column = col_max+1-column
  return(log(column))
}

#Remove outliers from retest data the way they have been from test data
retest_data = cbind(sub_id = retest_data$sub_id, as.data.frame(apply(retest_data[, -which(names(retest_data) %in% c("sub_id"))], 2, remove_outliers)))

#Transform necessary vars and rename those columns
for(i in 1:length(names(retest_data))){
  tmp_col = names(retest_data)[i]
  if(tmp_col %in% logTr_vars$var){
    sign = logTr_vars$sign[which(tmp_col == logTr_vars$var)]
      if(sign == 'positive'){
        retest_data[,tmp_col] = log(retest_data[,tmp_col])
        names(retest_data)[i] = paste0(tmp_col, '.logTr')
      }
      else if(sign == 'negative'){
        retest_data[,tmp_col] = neg_log(retest_data[,tmp_col])
        names(retest_data)[i] = paste0(tmp_col, '.ReflogTr')
      }
    }
}

```{r echo = FALSE}
rm(i, sign, tmp_col)
```

Check transformations are correct
```{r}
retest_logTr_vars = data.frame(grep('logTr', names(retest_data), value=T))
names(retest_logTr_vars) = c('var')
retest_logTr_vars$sign = ifelse(grepl(".ReflogTr",retest_logTr_vars$var), 'negative', 'positive')
retest_logTr_vars$var = gsub(".logTr|.ReflogTr", "", logTr_vars$var)
retest_logTr_vars == logTr_vars
```

```{r echo=FALSE}
rm(logTr_vars, retest_logTr_vars, meaningful_vars)
```

# Correlations for all cleaned meaningful variables
Spearman only

cor_df <- data.frame(spearman = rep(NA, length(matching_dv_columns)))

for(i in 1:length(matching_dv_columns)){
  cor_df[matching_dv_columns[i], 'spearman'] <- cor(retest_data[,matching_dv_columns[i]],retest_subs_test_data[,matching_dv_columns[i]], method = 'spearman', use = "pairwise.complete.obs")
}
```

Distribution of all Pearson and Spearman correlations
```{r warning=FALSE, message=FALSE}
cor_df %>%
  ggplot(aes(spearman))+
  geom_histogram()+
  theme_bw()+
  ggtitle('Distribution of Spearman correlations for cleaned meaningful variables')
```

Survey reliabilities are noticably higher than task reliabilities.
```{r warning=FALSE, message=FALSE}
cor_df %>%
  ggplot(aes(spearman))+
  geom_histogram()+
  theme_bw()+
  facet_wrap(~task)
```

Mean reliability and the number of variables for both the surveys and the tasks
```{r}
cor_df %>%
  group_by(task) %>%
  summarise(mean_spearman = mean(spearman),
            num_vars = n())
```

#### For tasks
```{r}
cor_df %>%
  filter(task == 'task') %>%
  select(dv, spearman) %>%
  arrange(-spearman)
```

- Demographics are excluded from this list of variables.  
- Holt and Laury test-retest is the worst for *surveys*. But note that this is not a survey that has been studied extensively psychometrically like others. This is more similar to out tasks (one that is used more often by economists). In previous data I have worked retest reliablities for this task was ~.35 so the results here are in line with this.  
- Otherwise the lowest survey reliabilities are 0.5.
```{r}
cor_df %>%
  filter(task == 'survey') %>%
  select(dv, spearman) %>%
  arrange(-spearman)
```

Keeping a record of the distributions of the cleaned meaningful variables.

```{r}
# Function to check if the variables that went in to cor_df have normal distributions
diff_from_normal <- function(col){
  return(shapiro.test(col)$p.value < 0.05)
}

# Create and organize new df where distribution of vars data is stored. First for retest data only; then for test data for subjects who have retest data - restricting here since these are the subsets of data that went in to the correlation matrix. Merge and cleanup.

dist_retest_data <- as.data.frame(apply(retest_data[,matching_dv_columns], 2, diff_from_normal))
names(dist_retest_data) <- c('retest_not_normal')
dist_retest_data$dv <- row.names(dist_retest_data)
row.names(dist_retest_data) <- seq(1:nrow(dist_retest_data))
dist_retest_data = dist_retest_data %>% select(dv, retest_not_normal)

rm(dist_retest_data, dist_retest_subs_test_data)

# Add whether dv's are tasks or survey to distribution dfs

dist_data = cor_df %>%
  select(dv, task) %>%
  left_join(dist_data, by = 'dv')
```

**Are the same vars that are not normally distributed in the retest data not normal in test data as well?**  
List of variables that have different distributions in the two time points.
```{r}
dist_data %>%
  filter(retest_not_normal != test_not_normal)
```

@@

We should probably take a closer look at model fits instead but this could be a first step.

The cleaned meaningful variables do not seem to include neither the EZ estimates nor accuracies
so analyses on this [here](https://htmlpreview.github.io/?https://github.com/zenkavi/SRO_Retest_Analyses/blob/master/SRO_Retest_Analyses.nb.html) are more informative. Otherwise just looking at this one might be tempted to conclude that rt's have higher reliabilities but clearly

```{r message=FALSE, warning=FALSE, fig.height=12, fig.width=9}
cor_df%>%
  filter(grepl('EZ|hddm|_rt|_acc', dv)) %>%
  mutate(var_type = ifelse(grepl('EZ',dv), 'EZ', ifelse(grepl('hddm',dv), 'hddm', 'raw')),
         var = ifelse(grepl('_rt', dv), 'rt', ifelse(grepl('_acc', dv), 'acc', ifelse(grepl('_non_decision', dv), 'non_decision', ifelse(grepl('_thresh', dv), 'thresh', ifelse(grepl('_drift', dv), 'drift', NA)))))) %>%
  ggplot(aes(spearman))+
  geom_histogram()+
  theme_bw()+
  facet_wrap(var_type~var)
```

```{r}
cor_df%>%
  filter(grepl('EZ|hddm|_rt|_acc', dv)) %>%
  mutate(var_type = ifelse(grepl('EZ',dv), 'EZ', ifelse(grepl('hddm',dv), 'hddm', 'raw')),
         var = ifelse(grepl('_rt', dv), 'rt', ifelse(grepl('_acc', dv), 'acc', ifelse(grepl('_non_decision', dv), 'non_decision', ifelse(grepl('_thresh', dv), 'thresh', ifelse(grepl('_drift', dv), 'drift', NA)))))) %>%
  group_by(var_type, var) %>%
  summarise(mean_spearman = mean(spearman, na.rm=T),
            sem_spear = sem(spearman),
            num_vars = sum(ifelse(is.na(spearman)==FALSE,1,0))) %>%
  arrange(var_type)
```

@@

First do a t-test on all measures of T1 for subjects with and without retest to confirm that these two samples do not differ from each other significantly (FDR corrected)
```{r}
ttest_df = data.frame(t_stat = rep(NA, length(matching_dv_columns)),
                      p_val_raw = rep(NA, length(matching_dv_columns)))

@@

  ttest_df[matching_dv_columns[i], 'p_val_raw'] <- t.test(retest_subs_test_data[,matching_dv_columns[i]], rest_subs_test_data[,matching_dv_columns[i]])$p.value
}

ttest_df$dv = row.names(ttest_df)
row.names(ttest_df) = seq(1:nrow(ttest_df))
ttest_df$task = 'task'
ttest_df[grep('survey', ttest_df$dv), 'task'] = 'survey'

#correct p-values 
ttest_df$p_val_fdr = p.adjust(ttest_df$p_val_raw, method='fdr')
```

```{r echo=TRUE}
rm(i)
```

Distribution of t statistics and p values

```{r warning=FALSE, message=FALSE}
ttest_df %>%
  gather(key, value, -dv, -task) %>%
  ggplot(aes(value)) +
  geom_histogram()+
  theme_bw()+
  facet_wrap(key~task, scales='free')
```

@@

13 variables have raw p values <0.05 BUT FDR corrected p values for these are not <0.05.

```{r}
ttest_df %>%
  filter(p_val_raw<0.05) %>%
  select(dv, task, t_stat, p_val_raw, p_val_fdr)
```

@@

Efa on the people you don't have retest on and then predict how well this holds up for people w retest on t1 and on t2.

I haven't consulted other factor analyses that have been done on the full dataset in doing this but will be happy to adjust as necessary.

```{r}
retest_data_mngf <- retest_data[,matching_dv_columns]
retest_subs_test_mngf <- retest_subs_test_data[,matching_dv_columns]
rest_subs_test_mngf <- rest_subs_test_data[,matching_dv_columns]
